

## Results of a small scale forest inventory

First, let's have a look at the results of 10 forest plots measured by the exploration crew. The data is stored in the object `fi_agb`. By running the object name in the console we will see the results:

```{r fi-demo}

fi_agb

```

*ADD variable description*

\  

The 10 plots come from a random sample in one of the Evergreen Forest of ```r newland_name```. The crew also shared the plot level aboveground biomass in the table name `fi_pagb`. 

**Your turn! Run the table `fi_pagb` in the console below to see the table basic information.**

```{r fi-cons, eval=F, class.source="Rchunk-cons"}

## Console
##

```


```{r fi-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

fi_pagb

```


*ADD variable description*

\  

Given that the crew chose the plot location randomly, we can use simple aggregating functions to re-calculate the forest mean aboveground biomass and confirm their calculation. We will use a mix of base R and `tidyverse` functions, in particular `summarise()`. This function is very handy to aggregate numerical variables in different columns of our plot table and can even be used in combination to `group_by()` to summarize numerical variables based on category variables of the same table (more on that later, with the full NFI data). 

Let's see for example how to calculate the mean basal area of the forest with its standard deviation, and save the results in a table `exfi_ba`. After creating the table, we run the table name in the console to display its content.

```{r fi-ba}

exfi_ba <- fi_pagb %>%
  summarise(
    mean_ba = mean(plot_ba),
    sd_ba   = sd(plot_ba)
  )

exfi_ba

```

\  

**Your turn! Create a table `exfi_agb` with the columns `mean_agb` and `sd_agb` and display the results.**

```{r fi-agb-cons, eval=F, class.source="Rchunk-cons"}

## Console
exfi_agb <- fi_pagb %>%
  summarise(
    mean_agb = mean(___),
    sd_agb   = sd(___)
  )

exfi_agb

```

```{r fi-agb-sol, class.source="Rchunk-sol", class.output="Rout-sol"}

exfi_agb <- fi_pagb %>%
  summarise(
    mean_agb = mean(plot_agb),
    sd_agb   = sd(plot_agb)
  )

exfi_agb

```

